Parallelization of Fast $\ell_1$-Minimization for Face Recognition

### 2011-02-18 paper/main.tex
fixed algorithm alm_rec and updated some numbers

The ALM algorithm requires many matrix-vector operations, this gives many-core
processors such as GPUs a computational advantage for large problem sizes.
GPUs are comprised of several streaming multiprocessors (SMP) (15 for our
NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups
of 32 threads called \emph{warps}.  These warps are then grouped into higher
levels called \emph{thread blocks} and \emph{grids}.

Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
For our system, each SMP has 64 kb of L1 cache with a total of 960 kb of L1 cache/shared memory and
768kb of L2 cache. However, this increase in performance does not come for
free. As the GPU has its own memory system, any data the GPU uses must be
transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
However, the cost of these transfers are amortized by doing large amounts of
computation at once, requiring code to be run completely on the GPU.  In 2011, NVIDIA is developing a new ARM-based CPU system, dubbed 
"Project Denver" to fix performance issues such as the limited bandwidth due to PCI-E \cite{Dally-Web}.

@@

\begin{algorithm}[t]
\caption{\bf (Augmented Lagrange Multiplier Method for
Recognition)}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\y \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \y$, $\blamda_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \textup{shrink}\left(\y - A\x_k +
\frac{1}{\mu}\blamda_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
\frac{1}{\gamma}A^T\left(\y - A\z_l - \e_{k+1} +
\frac{1}{\mu}\blamda_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\blamda_{k+1} \leftarrow \blamda_k + \mu (\y - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\label{alg:alm_rec} 
\end{algorithm}

@@

The first experiment is designed to compare the two implementations of PALM
that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm}).  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d/n$ varied from
$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
The ground truth signals, $x_0$ and $y_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.    We set the number of inner iterations at 50 and outer iterations at 5000.
We also varied the different inner and tolerance levels from 1e-3 to 1e-6.
We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.

@@

We pl

@@

We plot the resulted of the simulation benchmark below.  The x-axis represents the size of the 
$A$ and $A'A$ matrix and the y-axis represents the average amount of time it to complete
a single problem instance.

>>>>

The ALM algorithm requires many matrix-vector operations, this gives many-core
processors such as GPUs a computational advantage for large problem sizes.
GPUs are comprised of several streaming multiprocessors (SMP) (14 for our
NVIDIA Fermi) which each contains 32 floating-point units.  Each SMP contains
its own L1-cache and shared memory and all SMPs share a common L2-cache.  The
GPU executes programs as SIMT (Single Instruction Multiple Threads) in groups
of 32 threads called \emph{warps}.  These warps are then grouped into higher
levels called \emph{thread blocks} and \emph{grids}.

Fermi's inverted memory hierarchy (aggregate l1 capacity is greater than l2 capacity) is designed to support massively parallel workloads in which there is more task-private data than shared data, providing an increase in performance and a computational advantage over CPUs.
For our system, each SMP has 64 kb of L1 cache with a total of 896 kb of L1 cache/shared memory and
768kb of L2 cache. However, this increase in performance does not come for
free. As the GPU has its own memory system, any data the GPU uses must be
transferred from RAM to the GPU over PCI-Express, which has limited bandwidth.
However, the cost of these transfers are amortized by doing large amounts of
computation at once, requiring code to be run completely on the GPU.  In 2011, NVIDIA is developing a new ARM-based CPU system, dubbed 
"Project Denver" to fix performance issues such as the limited bandwidth due to PCI-E \cite{Dally-Web}.

@@

Below is the psuedocode solving the minimization problem posed in \eqref{eq:l1min_denoise}.

\begin{algorithm}[t]
\caption{\bf (Augmented Lagrange Multiplier Method for
Recognition)}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} $\bb \in \Re^m$, $A \in \Re^{m \times n}$,
$\x_1 = \mathbf{0}$, $\e_1 = \bb$, $\y_1 =
\mathbf{0}$.
\WHILE{not converged ($k = 1,2,\ldots$)}
\STATE $\e_{k+1} = \textup{shrink}\left(\bb - A\x_k +
\frac{1}{\mu}\y_k, \frac{1}{\mu}\right)$;
\STATE $t_1\leftarrow 1$, $\z_1 \leftarrow \x_k$, $\w_1 \leftarrow \x_k$;
\WHILE{not converged ($l = 1,2,\ldots$)}
\STATE $\w_{l+1} \leftarrow \textup{shrink}\left(\z_l +
\frac{1}{\gamma}A^T\left(\bb - A\z_l - \e_{k+1} +
\frac{1}{\mu}\y_k\right), \frac{1}{\mu\gamma}\right)$;
\STATE $t_{l+1} \leftarrow \frac{1}{2}\left( 1 +
\sqrt{1+4t_l^2}\right)$;
\STATE $\z_{l+1} \leftarrow \w_{l+1} + \frac{t_l - 1}{t_{l+1}}(\w_{l+1} - \w_l)$;
\ENDWHILE
\STATE $\x_{k+1} \leftarrow \w_{l}$,  \; $\y_{k+1} \leftarrow \y_k + \mu (\bb - A\x_{k+1} - \e_{k+1})$;
\ENDWHILE \STATE
{\bf Output:} $\x^* \leftarrow \x_k, \e^* \leftarrow \e_k$.
\end{small}
\end{algorithmic}
\label{alg:alm_rec} 
\end{algorithm}

@@

The first experiment is designed to compare the two implementations of PALM
that solve a single large instance of the $\ell_1$-minimization problem (Algorithm \ref{alg:alm_rec}).  
The measurement $d \times n$ matrix $A$ was generated as a random Gaussian matrix, with each entry
randomly generated from the standard normal and normalized with unit norm.  The ratio of $d:n$ varied from
$10:1$ to $10:5$ with the dimension of $d$ of matrix $A$ varying from 500 to 8000.
The ground truth signals, $x_0$ and $y_k$ had a sparsity rate of 0.1 with elements generated using the standard normal and normalized with unit norm.
The measurement vector was generated with the equation $\y_0 = A \x_0 + \y_k$.    We set the number of inner iterations at 50 and outer iterations at 5000.
We also varied the different inner and tolerance levels from 1e-3 to 1e-6.
We averaged the the time taken for the ALM algorithm to solve 10 instances of a problem.

@@

@@

We plot the resulted of the simulation benchmark below.  The x-axis represents the size of the 
$A$ and $A^TA$ matrix and the y-axis represents the average amount of time it to complete
a single problem instance.
